---
title: "Trilogy Use Case: Tricksters"
author: "Josh Hagedorn" 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    theme:
      version: 5
      bootswatch: litera
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse); library(plotly); library(tidytext)
# tmi <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/d57c2cefd0b216c8ce5c251f618c3e931c732d0a/data/tmi.csv")
# atu_df <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/d57c2cefd0b216c8ce5c251f618c3e931c732d0a/data/atu_df.csv")
# atu_seq <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/d57c2cefd0b216c8ce5c251f618c3e931c732d0a/data/atu_seq.csv")
# aft <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/d57c2cefd0b216c8ce5c251f618c3e931c732d0a/data/aft.csv")

tmi <- read_csv("../../trilogy/data/tmi.csv")
atu_df <- read_csv("../../trilogy/data/atu_df.csv")
atu_seq <- read_csv("../../trilogy/data/atu_seq.csv")
aft <- read_csv("../../trilogy/data/aft.csv")
```

# Use case

I'm in the midst of reading Lewis Hyde's *Trickster Makes the World*.

# Get trickster motifs

First, we can try a simple query using the word "trickster"

```{r}

trickster <- 
  tmi %>% 
  filter(str_detect(motif_name,regex("trickster", ignore_case = T))) %>%
  select(id:level)

```

Next, let's expand to look at any motif involving tricks.  We're going to use regular expressions (or, *regex*) to identify patterns in text, and we'll save these as a `list` object:

```{r}

regexes <- list()
regexes$trick <- "^trick | trick | trick$|^tricks | tricks | tricks$"

tricks <- 
  tmi %>% 
  filter(str_detect(motif_name,regex(regexes$trick,ignore_case = T)))

rm(trickster)
```

In Hyde's book, he identifies a number of other themes which accompany trickster tales, such as:

- __Bait.__ *Synonyms*: bait, trap, snare, catch, net
- __Stealing.__ *Synonyms*: steal, rob
- __Transformation.__ *Synonyms*: shape-shift, transform
- __Boundaries.__ *Synonyms*: border, boundary, cross-roads

We can make some regexes as search strategies, but keep in mind that these should be refined by looking at the data that's returned:

```{r}
regexes$bait <- "bait|trap|snare|catch| net "
regexes$steal <- "steal|rob"
regexes$transform <- "shape-shift|in.*shape|take.*shape|assume.*shape|transform"
regexes$border <- "border|boundary|boundaries|cross-roads"
```

We can add to these anytime (since I haven't finished the book yet), but let's start here and develop a subset of the `tmi` which contains trickster-related motifs.

```{r}
tricks <-
  tmi %>%
  select(id:chapter_id) %>%
  mutate(
    trickster = if_else(str_detect(motif_name,regex("trickster", ignore_case = T)),T,F),
    trick = if_else(str_detect(motif_name,regex(regexes$trick,ignore_case = T)),T,F),
    bait = if_else(str_detect(motif_name,regex(regexes$bait,ignore_case = T)),T,F),
    steal = if_else(str_detect(motif_name,regex(regexes$steal,ignore_case = T)),T,F),
    transform = if_else(str_detect(motif_name,regex(regexes$transform,ignore_case = T)),T,F),
    border = if_else(str_detect(motif_name,regex(regexes$border,ignore_case = T)),T,F)
  ) %>%
  rowwise() %>%
  mutate(any = sum(c_across(trickster:border))) %>%
  ungroup() %>%
  filter(any > 0)
```

# Find tales 

Joins: Get extended set of trickster tales

Identify all tail types, which include these motifs

Pull out metadata of sources and countries represented in the index

Check to see which values are highlighted in trickster tales, since these tales are amoral

