---
title: "Trilogy Workshop"
description: |
  An introduction to using data science tools for folklore taxonomy and text analysis
author:
  - name: Josh Hagedorn 
    url: https://www.j-hagedorn.com/
    orcid_id: 0000-0001-8026-7562
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    code_folding: true
---

# Load packages and trilogy data sets

Let's start by loading the materials we'll need:

- the R packages we'll use
- the `trilogy` datasets

The main package we'll use is the [tidyverse](https://www.tidyverse.org/), which is actually a collection of R packages with consistent design philosophy, grammar, and data structures.

To pull the current versions of the datasets, we'll follow the steps outlined in the [Getting Started > Use in Reproducible Research](https://github.com/j-hagedorn/trilogy/blob/master/docs/vignettes/getting_started.md#use-in-reproducible-research) vignette.  That's why you'll see a long, alphanumeric code in the links below, specifying precisely what version of the data is being used.

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

tmi <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/490a989d199d9440f9210bfe6ede797b03c4e26e/data/tmi.csv")
atu_df <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/490a989d199d9440f9210bfe6ede797b03c4e26e/data/atu_df.csv")
atu_seq <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/490a989d199d9440f9210bfe6ede797b03c4e26e/data/atu_seq.csv")
aft <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/490a989d199d9440f9210bfe6ede797b03c4e26e/data/aft.csv")

```

This allows us to explicitly reference a version of the data so that any research we do can be precisely replicated by others.  For instance, if you wanted to pull an old (and not yet cleaned-up) version of the `aft` dataset, you'd just need to go back in the GitHub history and run the following:

```{r old_aft, echo=TRUE}
old_aft <- read_csv("https://raw.githubusercontent.com/j-hagedorn/trilogy/f0fb12d108734847114f17980b05686a26305e38/data/aat.csv")
```


# Inspecting the data

## Motifs

## Tale Types

## Annotated Tales

# 

- Joining together data sets by primary key
- First analysis: coverage of motif index in the tale type index
- Visualization
- Adding citations
- Where to go next?



